% =============================================================================
% From Eclipse Demon to Lunar Apogee
% Companion paper — pdflatex + CJKutf8, two-column
% =============================================================================
\documentclass[10pt,a4paper]{article}

% --- Packages (matches when-the-stars-disagree.tex) -------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{CJKutf8}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[margin=0.75in]{geometry}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{longtable}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{xurl}
\usepackage{authblk}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{gensymb}
\usepackage{adjustbox}
\usepackage[framemethod=tikz]{mdframed}

\pgfplotsset{compat=1.18}

\setlength{\columnsep}{18pt}
\setlength{\emergencystretch}{3em}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=blue!60!black,
}

% CJK helper
\newcommand{\zh}[1]{\begin{CJK}{UTF8}{bsmi}#1\end{CJK}}

% ORCID helper
\newcommand{\orcidlink}[1]{%
  \href{https://orcid.org/#1}{\textcolor{green!50!black}{\textsf{iD}}}%
}

% pgfplots colors
\definecolor{apogeeblue}{HTML}{2166AC}
\definecolor{nodered}{HTML}{D6604D}

% --- Title -------------------------------------------------------------------

\title{%
  \textbf{From Eclipse Demon to Lunar Apogee:}\\[4pt]
  \large Six Modes of Transformation in Cross-Civilizational\\
  Astronomical Transmission%
}

\author[1]{Albert Kwun Tai Hui\,\orcidlink{0009-0000-8978-0231}}
\affil[1]{Security Ronin}

\date{March 2026}

% =============================================================================
\begin{document}

\twocolumn[
\maketitle

\begin{abstract}
When astronomical concepts travel across civilizational boundaries, they do not
arrive unchanged. This paper argues that such transformations are systematic
rather than random, and proposes a six-mode theoretical framework---transliteration,
reidentification, indigenous restructuring, hybridization, foreign correction, and
algorithmic fossilization---derived from an extended case study spanning thirteen
centuries.

The case study follows the Indian shadow planet Ketu (\zh{計都}) and its coordinate
reference frame as both were absorbed into Chinese \zh{七政四餘} (Seven Governors,
Four Remainders) astrology.  On the coordinate axis the paper traces four
successive regime changes: ecliptic sidereal coordinates in the Song dynasty
(\zh{三辰通載}); equatorial sidereal in the Yuan (\zh{鄭氏星案}); a novel
quasi-ecliptic sidereal hybrid in the Ming (\zh{果老星宗}); and ecliptic tropical
under the Jesuit reform of 1645.  In parallel, Ketu's astronomical identity was
quietly reidentified---from the descending lunar node to the osculating lunar
apogee---a shift demonstrated by Niu Weixing's 1994 recomputation of Tang-dynasty
ephemeris data \cite{niu1995ketu}, before being reverted by Schall von Bell in the
\textit{Shixian} calendar.

The paper contributes: (1)~computational verification of the apsidal identification
using modern JPL ephemeris algorithms; (2)~the first rigorous definition of the
\zh{似黃道} (quasi-ecliptic) coordinate system; and (3)~the six-mode framework as
a reusable instrument for the comparative history of transmitted astronomy.
\end{abstract}
\vspace{12pt}
]

% === §1 Introduction =========================================================
\section{Introduction}\label{sec:intro}
% ~1,500 words — Task 6

\subsection*{The Silk Road as a Channel for Astronomical Ideas}

Astronomical knowledge traveled the Silk Road alongside silk, spices, and
religious texts.  Between the second and ninth centuries~CE, Greek planetary
models, Indian \textit{siddh\=anta} tables, Persian \textit{z\={\i}j} manuals,
and Babylonian omen compendia all made their way into Central Asia and China via
a network of court astronomers, Buddhist missionaries, Sogdian merchants, and
imperial tribute missions.  Unlike a bolt of silk, however, an astronomical concept
does not arrive at its destination intact.  It is translated into a new language,
fitted into an existing institutional structure, combined with native techniques,
challenged by fresh observations, and eventually enshrined in software whose users
have forgotten what it ever meant.  These transformations leave detectable traces
in historical texts---and in the computed positions of celestial bodies.

This paper argues that the transformations are systematic rather than random.
Using a single case study that spans thirteen centuries and four Chinese dynasties,
we identify six recurring patterns---six \emph{modes} of transformation---that
structured the reception of Indian astronomy in China.  The case study concerns
two related objects: the Indian shadow planet Ketu (\zh{計都}) and the coordinate
reference frame used to specify its position.  Both entered Chinese astronomy in
718~CE through the translation of the \textit{Jiuzhi li} (\zh{九執曆},
\textit{Nine Luminaries Calendar}) by Gautama Siddh\=artha (\zh{瞿曇悉達}), a
scholar of Indian descent serving at the Tang court.  Both underwent a remarkable
sequence of transformations over the following millennium, and both continue to
influence popular Chinese-language astrology software today.

\subsection*{Two Parallel Threads}

The paper traces two intertwined threads that run from 718~CE to the present.

\textit{Thread~1: The coordinate reference frame.}  Every astronomical position
requires a reference frame---a choice of origin, fundamental plane, and direction
of increasing longitude.  The original Indian system used ecliptic coordinates
measured from a sidereal zero point (a fixed star rather than the moving
vernal equinox).  Chinese astronomy in the Tang and Song dynasties preferred
equatorial coordinates referenced to the twenty-eight lunar lodges (\zh{二十八宿},
\textit{xiu}).  When Ketu entered the Chinese tradition, a tension arose between
the imported ecliptic frame and the native equatorial one.  That tension was
resolved---and re-resolved---four times:
\begin{enumerate}[leftmargin=*, label=(\arabic*)]
  \item \textbf{Song ecliptic sidereal.}  The \zh{三辰通載} (c.~11th century~CE)
    presents Ketu positions in a coordinate frame that is recognisably ecliptic
    and sidereal, close to the Indian original.
  \item \textbf{Yuan equatorial sidereal.}  The \zh{鄭氏星案} (late Yuan, c.~14th
    century) converts all positions into equatorial coordinates aligned with the
    native lodge system---a thoroughgoing indigenization of the imported framework.
  \item \textbf{Ming quasi-ecliptic sidereal.}  The \zh{果老星宗} (Ming dynasty,
    c.~15th--16th century) introduces what we term the \zh{似黃道} (quasi-ecliptic)
    frame: a hybrid that applies ecliptic-style longitude increments along a plane
    that is neither the true ecliptic nor the equator.  This is the least-studied
    coordinate system in Chinese historical astronomy, and the paper offers the
    first formal definition.
  \item \textbf{Qing ecliptic tropical.}  The Jesuit-led calendar reform of 1645
    (\textit{Shixian li}, \zh{時憲曆}) reimposed European tropical ecliptic
    coordinates, resetting the accumulated drift and discarding the Ming hybrid.
\end{enumerate}
The trajectory is not a smooth evolution but a sequence of reversals---``there
and back again, but not quite'': the Qing frame resembles the Song one in being
ecliptic, but has shifted from sidereal to tropical reckoning.

\textit{Thread~2: The identity of Ketu.}  In Indian astronomy Ketu (\textit{ketu},
literally ``comet'' or ``banner'') designates the descending lunar node---one of
the two points where the Moon's orbit crosses the ecliptic.  When the Moon crosses
the descending node moving southward, conditions favour a lunar eclipse; when
it crosses the ascending node Rahu (\zh{羅睺}), conditions favour a solar eclipse.
The two nodes are therefore deeply embedded in eclipse prediction.

Niu Weixing's landmark 1994 analysis of the \zh{七曜攘災訣} (806~CE), a Tang
ephemeris attributed to the monk Amoghavajra's circle, demonstrated that the
Ketu positions recorded therein do \emph{not} track the descending node
\cite{niu1995ketu}.  They track the osculating lunar apogee---the point in the
Moon's elliptical orbit farthest from Earth.  The apogee was known to Indian
astronomers as \textit{mand\=occa} (``slow apex'') and in Persian-influenced
sources as \textit{K\=etus}; its conflation with the node appears to have
occurred during or shortly after the Tang transmission.  By the Ming period,
Ketu as apogee was thoroughly assimilated into \zh{七政四餘} practice, where it
joined Rahu (\zh{羅睺}), the Purple Star (\zh{紫氣}), and the Monthly Comet
(\zh{月孛}) as the four ``remainders''---invisible bodies that governed fate
alongside the seven visible planets.  The Jesuit reform then reverted Ketu to
the descending node, erasing five centuries of accumulated reidentification.

\subsection*{Six Modes of Transformation}

The six modes proposed here are not a priori categories imposed on the data.
They emerge from the case study as the recurring patterns by which astronomical
concepts were altered during transmission.  We state them here in condensed form;
each receives fuller treatment in the sections that follow.

\textbf{Mode~1: Transliteration.}  Foreign technical terms are adopted into the
receiving tradition with minimal semantic change, using phonetic approximation.
\textit{R\=ahu} becomes \zh{羅睺} (\textit{lu\'ohou}); \textit{Ketu} becomes
\zh{計都} (\textit{j\`{\i}d\=u}).  The sounds are borrowed; the astronomical
meanings---initially---travel with them.  Transliteration is the entry mechanism:
it creates a named slot in the receiving system before that slot has been filled
with indigenous content.

\textbf{Mode~2: Reidentification.}  The name persists while the astronomical
referent shifts.  The transformation is invisible at the lexical level but
detectable by comparing computed positions against historical ephemeris records.
Ketu's drift from descending node to osculating apogee is the central instance;
we suspect the shift happened in stages between 718 and 806~CE as Tang
practitioners noticed that the imported node-based algorithms gave increasingly
poor results and sought a better-fitting celestial body.

\textbf{Mode~3: Indigenous restructuring.}  The native tradition actively
overwrites the imported framework by substituting its own coordinate system,
calendar reckoning, or theoretical structure.  The Yuan equatorial conversion
is the clearest example: Yuan astronomers did not argue with the Indian ecliptic
frame; they simply replaced it with the equatorial lodge-coordinate system that
their \textit{Shoushi li} (\zh{授時曆}) reform had just made canonical.

\textbf{Mode~4: Hybridization.}  Elements of the imported and native systems are
combined into a novel structure that belongs fully to neither parent tradition.
The Ming \zh{似黃道} system is the prime instance.  It preserves ecliptic-style
longitude from the Indian source while anchoring the zero-point and plane to a
native star-count system; the result is not Indian ecliptic astronomy and not
Chinese lodge astronomy, but a tertium quid that persists in Chinese-language
software to this day.

\textbf{Mode~5: Foreign correction.}  An external authority---a new wave of
transmission, a conquering dynasty, or an imported expertise---imposes a reset
that overrides accumulated indigenous development.  Schall von Bell's 1645
\textit{Shixian li} reform is the paradigm case.  Backed by Qing imperial
authority, it discarded the Ming quasi-ecliptic frame, reverted Ketu to the
descending node, and replaced sidereal with tropical reckoning---all in a single
edict.

\textbf{Mode~6: Algorithmic fossilization.}  A living astronomical practice,
once encoded in tables or (later) software, becomes separated from its theoretical
foundation.  Users apply the algorithm without access to---or interest in---the
historical reasoning behind it.  Errors accumulate unnoticed because the algorithm
is self-consistent even if its foundations have shifted.  This mode is not
historically confined to the past: modern Chinese-language astrology software
routinely inherits the Ming \zh{似黃道} frame without labelling it as such,
producing systematic positional errors of several degrees when compared against
modern ephemerides.

\subsection*{Method}

The paper combines close reading of primary astronomical texts with computational
verification using modern ephemeris algorithms.  For the computational component
we use the Swiss Ephemeris library (version~2.10), which implements the JPL
DE441 planetary and lunar solution.  This allows us to ask a precise empirical
question: does a given historical Ketu ephemeris track the lunar apogee, the
descending node, or something else?  We convert historical sexagesimal positions
to modern decimal degrees, apply the appropriate calendar and epoch corrections,
and compare against Swiss Ephemeris output for the descending node and for the
osculating apogee at the same Universal Time.  The method follows and extends
the approach used by Niu Weixing \cite{niu1995ketu} for the \zh{七曜攘災訣} and
applied to Ming texts by Niu in his later synthesis \cite{niu2010fourremains}.
Computational details and residual tables are presented in \S\ref{sec:computation}.

\subsection*{Relationship to Existing Scholarship}

This paper builds on four converging research programmes.

Niu Weixing's identification of Ketu as the lunar apogee
\cite{niu1995ketu,niu2010fourremains} is the single most important prior result
for our argument.  Without Niu's recomputation, the reidentification mode would
remain invisible.  Our contribution is to situate Niu's finding within a broader
transformative sequence and to provide independent computational confirmation
using a modern ephemeris unavailable to Niu in 1994.

Bill Mak's work on pseudo-planets traces the historical trajectory of Rahu and
Ketu from Hellenistic and Indian origins into China and Japan, with particular
attention to Persian-mediated transmission \cite{mak2014pseudoplanets,mak2024persian}.
Mak documents the textual genealogy that brought shadow-planet concepts to East
Asia; the present paper takes up where Mak's textual history ends, asking what
happened to those concepts inside China over the long Ming--Qing transition.

Jeffrey Kotyk has produced detailed studies of the sinicization of Indo-Iranian
astrology in the Tang and Song courts \cite{kotyk2018sinicization,kotyk2017buddhist}.
Kotyk establishes the Buddhist monastic networks---particularly Amoghavajra's
school---through which Indian astronomical texts circulated in the eighth century.
Our \S\ref{sec:tang} on the Tang transmission depends heavily on Kotyk's
reconstruction of institutional context.

Liu Ts'un-yan's work on the Jesuit encounter with Chinese astrology
\cite{liu2020missionary} provides the documentary foundation for our account of
Mode~5 (foreign correction) in \S\ref{sec:jesuit}.  The Jesuit reform was not
merely a technical event; it was an act of epistemic authority, and Liu's analysis
of Schall von Bell's rhetorical strategies illuminates why the reversion of Ketu
to the descending node was politically possible in 1645.

A companion paper, ``When the Stars Disagree: A Cross-Civilizational Analysis
of Computational Errors in Astronomical Divination,'' presents a broader taxonomy
of error types arising from coordinate-frame mismatches across the Chinese,
Hellenistic, Indian, and Persian traditions.  The present paper concentrates on
the transformation \emph{mechanisms} rather than the error \emph{outcomes}.

\subsection*{Structure of the Paper}

Section~\ref{sec:indian} describes the Indian original: the conceptual status of
Rahu and Ketu in Sanskrit \textit{siddh\=anta} astronomy and the ecliptic sidereal
coordinate frame in which they were embedded.  Sections~\ref{sec:tang}--\ref{sec:jesuit}
trace the four dynasties and four coordinate regimes in sequence.
Section~\ref{sec:fossil} analyses the ongoing algorithmic fossilization in modern
software.  Section~\ref{sec:computation} presents the computational verification.
Section~\ref{sec:discussion} draws the six modes together into the theoretical
framework and discusses their applicability beyond the Ketu case.
Section~\ref{sec:conclusion} concludes.

% === §2 The Indian Original ===================================================
\section{The Indian Original: Rahu, Ketu, and Ecliptic Sidereal Coordinates}
\label{sec:indian}
% ~1,000 words — Task 7

\subsection*{Ecliptic Sidereal Coordinates}

Indian \textit{Jyotish} astronomy rests on a coordinate framework that differs
from both the Greek tropical system and the native Chinese equatorial tradition.
The fundamental reference plane is the \emph{ecliptic}---the apparent annual path
of the Sun against the background of fixed stars.  Along this plane the sky is
divided into twelve \textit{r\=asi} (zodiac signs), each occupying exactly 30° of
arc, yielding the familiar circle of 360°.  What distinguishes the Indian system
from its Greek counterpart is the anchor point: where Hellenistic and later
Western astronomy measures longitude from the vernal equinox (a moving point
that drifts westward through the stars at roughly 50 arcseconds per year due to
the precession of Earth's rotation axis), Indian astronomy measures from fixed
stars.  A position is therefore \emph{sidereal}---\textit{nirayana}, ``without
the \textit{ayana}'' or solstitial shift---rather than tropical
(\textit{s\=ayana}).

The distinction is consequential.  Over the roughly two millennia since the
Indian and Hellenistic systems last agreed on the position of the vernal equinox
(approximately 285~CE by the most widely used \textit{Lahiri} \textit{ayanamsha},
though estimates range from 285 to 560~CE depending on the \textit{ayanamsha}),
precession has opened a gap of nearly 24° between tropical and sidereal
longitudes.  A planet at 10° tropical Aries sits near 16° sidereal Pisces.
Every translation between Indian and Chinese sources must contend with this
offset \cite{pingree1981jyotih}.

The twelve \textit{r\=asi}---\textit{Mesha} (Aries) through \textit{M\={\i}na}
(Pisces)---serve as the containers within which planetary positions are reported.
They are equal divisions, not constellations of variable extent, so the
\textit{r\=asi} framework is a clean mathematical abstraction.  Each
\textit{r\=asi} is further subdivided into \textit{nakshatra} (lunar mansion)
segments, linking the zodiacal and lunar-mansion systems in a unified
architecture.  It was this architecture---ecliptic plane, 360° circle,
equal-division signs, nine celestial bodies---that Tang dynasty translators would
import wholesale into China in 718~CE.

\subsection*{Rahu and Ketu: The Eclipse Demons}

Within the nine \textit{graha} (``seizers'' or celestial influences) of Indian
astronomy, two occupy a peculiar metaphysical status.  Rahu and Ketu are not
visible planets but mathematical points: the two nodes where the Moon's orbital
plane intersects the ecliptic.  At these intersections, and only at these
intersections, solar and lunar eclipses are geometrically possible.  Because
eclipses were among the most feared celestial events in ancient societies, the
nodes acquired a mythology commensurate with their power.

The origin story is preserved in the \textit{Mah\=abh\=arata} and the
\textit{Bh\=agavata Pur\=ana}.  During the churning of the cosmic ocean the gods
produced \textit{am\d{r}ta}, the nectar of immortality.  A demon named
Svarbh\=anu disguised himself among the gods and drank from the vessel.  The Sun
and Moon recognized the impostor and alerted Vishnu, who immediately severed
Svarbh\=anu's head from his body with his discus.  Because the demon had already
swallowed the nectar, however, both head and tail survived as immortal fragments:
the head became Rahu, the tail became Ketu.  Eternally enraged, they periodically
swallow the Sun and Moon---producing eclipses---only for the luminaries to emerge
from the severed neck moments later.

This mythological account maps precisely onto the astronomical reality.  Rahu,
the ascending (north) lunar node, is where the Moon crosses from south to north
of the ecliptic; Ketu, the descending (south) node, is where it crosses from
north to south.  The two nodes are always exactly 180° apart---head and tail of
the same orbital intersection---and they precess westward around the ecliptic
with a period of approximately 18.6 years.  The identification is astronomically
unambiguous and remained stable across Indian tradition for more than a millennium
\cite{pingree1981jyotih}.

One complication deserves note before this Indian baseline is left behind.  The
Sanskrit word \textit{ketu} originally meant ``banner,'' ``bright appearance,''
or ``radiant body,'' and in earlier Vedic and post-Vedic literature it referred
primarily to comets \cite[p.~47]{kotyk2018sinicization}.  The identification of
\textit{ketu} specifically with the descending lunar node is a later development,
consolidated in the \textit{siddh\=anta} tradition but never entirely displacing
the older, broader sense.  This semantic flexibility---a name that had already
migrated once, from comet to node---prefigures the second migration that would
occur in Tang dynasty China, when \zh{計都} moved from node to apogee.

\begin{table*}[ht]
\centering
\small
\caption{Rahu and Ketu: Indian original vs.\ mature Chinese identification.}
\label{tab:rahu-ketu}
\begin{tabular}{@{}lll@{}}
\toprule
Body & Indian (\textit{Jyotish}) & Chinese (\zh{果老星宗}) \\
\midrule
\zh{羅睺} (Rahu) & Ascending node & Ascending node \\
\zh{計都} (Ketu) & Descending node & Osculating lunar apogee \\
\zh{月孛} (Yuebei) & --- & Mean lunar apogee \\
\zh{紫氣} (Ziqi) & --- & Fictitious ($\sim$28-year cycle) \\
\bottomrule
\end{tabular}
\end{table*}

\subsection*{The Clarity That Makes Transformation Visible}

The Indian system's identification of Rahu and Ketu is admirably
unambiguous. Two nodes, always opposite, always moving at the same rate,
always causally connected to eclipses: the conceptual architecture is
self-consistent and empirically grounded.  Ketu is the descending node,
full stop.  There is no competing function for it to perform, no
ambiguity about its period or direction of motion.

Table~\ref{tab:rahu-ketu} places this Indian baseline alongside the
mature Chinese identification that will be traced through the following
sections.  The contrast is arresting.  In the Chinese \zh{七政四餘}
system, \zh{羅睺} (Rahu) retains its Indian identity as the ascending
node, but \zh{計都} has been detached from the descending node
altogether and reattached to the osculating lunar apogee---a point that
moves in the \emph{same} direction as the Moon rather than opposite to
it, has a period of roughly 8.85 years rather than 18.6 years, and has
no direct causal relationship to eclipses.  A new body, \zh{月孛}
(\textit{Yueb\`ei}, ``lunar comet''), has been invented to track the
mean apogee, and a fourth fictitious point, \zh{紫氣} (\textit{Ziq\`{\i}},
``purple vapour''), has been added with a $\sim$28-year cycle that has
no Indian precedent.

This proliferation did not happen overnight.  It was the cumulative
product of successive transformations---transliteration, reidentification,
indigenous restructuring, and hybridization---that will be traced
dynasty by dynasty, beginning with the first moment of contact in
718~CE.

% === §3 The Tang Transmission =================================================
\section{The Tang Transmission: 718--806~CE}\label{sec:tang}
% ~1,500 words — Task 7

\subsection*{Transformation Mode~1: Transliteration}

The first mode of transformation is the simplest: foreign astronomical
concepts are adopted with minimal modification.  Names are transliterated,
tables are copied, procedures are translated as literally as the target
language permits.  The receiving culture does not yet possess the
institutional confidence to reshape what it has received; it is still
in the posture of a student before a prestigious foreign tradition.
This is the characteristic mode of the early Tang reception of Indian
astronomy.

\subsection*{The \textit{Jiuzhi li} of 718~CE}

The pivotal document is the \textit{Jiuzhi li} \zh{九執曆}
(\textit{Navagraha-karana}, ``Calendar of the Nine Seizers''), translated
in 718~CE by Gautama Siddh\=artha (\zh{瞿曇悉達}), an astronomer of
Indian descent whose family had settled in Chang'an over the preceding
generation.  Gautama served in the Tang Bureau of Astronomy
(\zh{太史局}) and brought to his translation both fluency in Sanskrit
astronomical texts and direct access to the imperial court's demand for
accurate calendrical and astrological computation
\cite{yabuuchi1989sui,kotyk2018sinicization}.

The \textit{Jiuzhi li} introduced into China a package of Indian
astronomical innovations that had no precedent in the native tradition.
Most consequential for positional astronomy was the adoption of a
360-degree circle, replacing the Chinese convention of dividing the
ecliptic into 365.25 degrees (one degree per day of the solar year).
Alongside this came decimal notation for fractional positions, sine
tables computed at 3.75° intervals (a characteristic feature of the
Indian \textit{siddh\=anta} tradition that derives ultimately from
Hellenistic chord tables), and epicyclic planetary models for computing
the true positions of the five visible planets from their mean motions.

Among the nine \textit{graha} that give the text its name, two were
purely mathematical points: Rahu (\zh{羅睺}) and Ketu (\zh{計都}).
Their Chinese names are phonetic transliterations---\zh{羅睺} for
``Rahu'' and \zh{計都} for ``Ketu''---the clearest possible signal that
the Tang astronomers were importing the concept as a foreign package
rather than mapping it onto any pre-existing Chinese category.  Both
entered Chinese astronomical vocabulary simultaneously, both designated
lunar nodes, and both were assigned the same westward precessional
motion of approximately 3.06° per month
\cite{yabuuchi1989sui,kotyk2018sinicization}.

The \textit{Jiuzhi li} also introduced the ecliptic as an explicit
reference plane.  Chinese astronomers had for centuries measured
celestial positions in equatorial coordinates: right ascension expressed
as entry into one of the twenty-eight lunar mansions (\zh{入宿度},
\textit{r\`uxiud\`u}) and polar distance (\zh{去極度}, \textit{q\`uj\'ud\`u}).
The ecliptic was not ignored in Chinese astronomy---the twelve
\textit{ci} (\zh{次}) had long served as a solar-position scheme---but
it was not the primary framework for planetary positions.  The
\textit{Jiuzhi li} brought with it a fully ecliptic-centric positional
system, creating an immediate tension between the imported framework and
the native one that would not be resolved for nearly six centuries.

\subsection*{The Iranian Intermediary}

The Tang transmission was not, however, a clean direct line from India
to China.  Kotyk's detailed philological analysis demonstrates that the
\textit{Jiuzhi li} and the astrological texts that accompanied it show
unmistakable signs of Persian and Sasanian intermediary influence
\cite{kotyk2018sinicization,kotyk2017buddhist}.  Calculation procedures,
parameter values, and terminology in several passages align more closely
with Persian \textit{z\={\i}j} tradition than with the Indian
\textit{siddh\=anta} sources from which the system ultimately derives.

Mak's recent study of the Persian astronomical channel into Tang China
strengthens this picture considerably \cite{mak2024persian}.  The
Sasanian empire maintained sophisticated astronomical institutions, and
Persian court astronomers had been assimilating and adapting Indian
methods for at least two centuries before the Tang transmission.  When
Indian astronomical ideas arrived in Chang'an, they had already been
filtered through a Persian cultural lens that introduced modifications---
in nomenclature, in parameter selection, in the relative prominence
accorded different celestial bodies---that are not always easy to
disentangle from the original Indian substrate.

This Iranian intermediary matters for the history of Ketu's
transformation.  If modifications to the node--apogee identification
were already under way in Persian astrology, the Chinese reidentification
may reflect not a purely Chinese innovation but the endpoint of a
gradual drift that began somewhere along the Silk Road between India
and Chang'an.  The evidence is not yet conclusive, but the Persian
channel must remain in view as a possible locus of the first
reidentification.

\subsection*{The \zh{七曜攘災訣} and the Apogee Identification (806~CE)}

The second key Tang text is the \textit{Qiy\`ao r\`angz\=ai ju\'e}
\zh{七曜攘災訣} (``Formulae for Averting Disasters Caused by the Seven
Luminaries''), a Buddhist astrological manual compiled by a
Br\=ahma\d{n}a priest from western India sometime after 806~CE, the
epoch year (Year~1 of the Yuanhe \zh{元和} reign period) embedded in its
planetary tables.  The text would not survive in Chinese repositories;
it was brought to Japan in 865~CE by the Japanese monk Shuy\=u
(\zh{宗叡}) and preserved within the \textit{Sukuy\=od\=o}
(\zh{宿曜道}, ``Way of the Lunar Mansions and Stars'') tradition, where
it remained effectively unknown to Chinese and Western scholarship until
modern times \cite{kotyk2017buddhist}.

The \textit{Qiy\=ao r\`angz\=ai ju\'e} contains multi-decade ephemerides for
all nine \textit{graha}, including a Rahu table spanning 93~years and a
Ketu table spanning 62~years.  In 1994, the Chinese historian of science
Niu Weixing subjected the Ketu ephemeris to a systematic computational
reanalysis, comparing its tabulated positions against two candidates:
the descending lunar node as defined in Indian \textit{siddh\=anta}
astronomy, and the osculating lunar apogee \cite{niu1995ketu}.  The
result was unambiguous: the Ketu ephemeris fits the motion of the
\emph{apogee}, not the node.  The period, the direction of motion, and
the epoch position all correspond to the lunar apogee; none correspond
to the descending node.

This finding establishes a terminus ante quem for the reidentification.
By 806~CE---within roughly one century of the \textit{Jiuzhi li}
translation---\zh{計都} had already ceased to designate the descending
lunar node and had been reassigned to track the osculating apogee.  The
name persisted; the referent changed completely.

\subsection*{Coordinate Thread: The Ecliptic in an Equatorial World}

The coordinate dimension of the Tang transmission is a story of
institutional coexistence rather than replacement.  The \textit{Jiuzhi li}
introduced an ecliptic framework, but the Bureau of Astronomy continued
to produce the official calendar (\textit{li} \zh{曆}) in the native
equatorial system.  Indian ecliptic methods were used for astrological
purposes---planetary hour lords, eclipse prediction, the \textit{graha}
positions in horoscopes---while the equatorial system retained its
authority for the state rituals tied to the solar terms and the
twenty-eight lunar mansions.

The result was a bifurcated institutional landscape.  Buddhist
astrologers working from Indian and Persian sources operated in ecliptic
coordinates; the official Bureau computed in equatorial ones.  The two
systems were not formally reconciled, and conversion procedures between
them were approximate at best.  This bifurcation, established in the
early eighth century, would persist as an unresolved tension through the
Song dynasty, when the first serious attempt at a unified ecliptic
sidereal system for Chinese \textit{qizheng siyu} (\zh{七政四餘},
Seven Governors and Four Remainders) astrology would finally be made.

\subsection*{Identity Thread: Mode~2 Already in Motion}

The \textit{Qiy\=ao r\`angz\=ai ju\'e}'s apogee ephemeris means that by
the end of the Tang period, both transformation modes are already
visible.  Mode~1 (transliteration) is represented by the \textit{Jiuzhi
li}: foreign names and methods are adopted wholesale, with \zh{羅睺}
and \zh{計都} entering Chinese as phonetic loan words for Indian
concepts.  Mode~2 (reidentification) is already under way: the name
\zh{計都} has been retained but silently detached from its original
referent.

What is striking about this early reidentification is its lack of
documentary justification.  No Tang text known to survive explains
\emph{why} Ketu should track the apogee rather than the node.  The
change is not announced, not argued for, not defended.  It appears in
the ephemeris tables as a fait accompli, suggesting that it may have
occurred upstream---in the Persian intermediary tradition, or in a
lost stratum of Indian regional astronomy---rather than being a
conscious Chinese innovation.  Whatever its origin, the reidentification
was consequential: it would shape every subsequent Chinese treatment of
the Four Remainders, creating the distinctive asymmetry---Rahu as node,
Ketu as apogee---that defines the mature \zh{七政四餘} system and
persists in software to this day.

% === §4 The Song System =======================================================
\section{Ecliptic Sidereal: The Song System (\zh{三辰通載})}
\label{sec:song}
% ~1,200 words — Task 8

\subsection*{Transformation Mode~2: The Post-Transliteration State}

The Tang transmission planted Indian astronomical seeds in Chinese soil.
The Song dynasty (960--1279~CE) represents the first flowering: a codified,
institutionalized system that preserved the imported framework with remarkable
fidelity.  Before the indigenous restructuring of the Yuan and the Jesuit
correction of the Qing, there existed a Chinese \zh{七政四餘} practice that
looked, structurally, very much like what a medieval Indian astrologer would
recognize.  The \zh{三辰通載} (\textit{S\={a}nch\'{e}n t\={o}ngz\`ai},
``Comprehensive Record of the Three Luminaries'') is our primary evidence for
this baseline state.

\subsection*{The \zh{三辰通載} as Earliest Codification}

The \zh{三辰通載} is the earliest extant Chinese text to codify the full
\zh{七政四餘} system in a form recognisably continuous with its Indian
antecedents.\footnote{Direct access to the \zh{三辰通載} text remains limited;
our characterization relies on secondary analyses, principally
\citealt{ho2003astrology} and \citealt{mak2014pseudoplanets}.}
Attributed to the Song period and concerned with what Chinese tradition called
the ``three luminaries''---Sun, Moon, and stars---the text assembles the
seven visible planets (\zh{七政}: Sun, Moon, Mercury, Venus, Mars, Jupiter,
Saturn) together with the four ``remainders'' (\zh{四餘}) into a unified
predictive scheme.

What makes the \zh{三辰通載} a baseline rather than merely an early text
is precisely what it does \emph{not} do.  It does not translate the imported
ecliptic frame into native equatorial coordinates.  It does not recast
planetary dignities in terms of the Five Phases (\zh{五行}).  It does not
subject the pseudo-planets to a programme of sinicization.  The system
is cosmopolitan in the most literal sense: it operates within coordinate
conventions and interpretive categories that arrived from the west and
were not yet remade into something distinctively Chinese.

\subsection*{Coordinate Thread: The Ecliptic Sidereal Frame}

The \zh{三辰通載} employs what we designate the \zh{黃道恆星製}
(\textit{hu\'{a}ngd\`ao h\'{e}ngx\=ing zh\`{\i}}, ``ecliptic sidereal
system'').  Planetary longitudes are measured along the ecliptic from a
fixed sidereal zero-point---the same fundamental choice that underpins
Sanskrit \textit{siddh\={a}nta} astronomy.  The twelve zodiacal houses
(\zh{十二宮}) are equal 30\textdegree{} divisions of the ecliptic, each
mapped to one of the twelve Earthly Branches (\zh{地支}): \zh{子} (Zi)
for Aquarius, \zh{丑} (Chou) for Capricorn, and so on in the familiar
sequence.  This direct correspondence between the Indian zodiac and the
Chinese duodenary cycle is not a coincidence; it is a deliberate
translation, executed during the Tang transmission, that anchored the
foreign framework to an existing Chinese indexing system
\citep{kotyk2018sinicization}.

Within each house, the interpretive machinery is recognisably Jyotish.
Aspects (\zh{刑沖合會}) correspond to the Indian \textit{dr\d{s}\d{t}i}
doctrine.  Planetary dignity---exaltation, detriment, and the like---maps
to \textit{ucca} and \textit{n\={\i}ca} as transmitted through Persian
intermediaries.  Crucially, no Five-Phase (\zh{五行}) overlay is present.
The planets are not assigned to Wood, Fire, Earth, Metal, and Water in
their astrological capacities; they remain the seven Greco-Indian wanderers,
differentiated by their behavior within the ecliptic frame rather than by
their membership in a Chinese cosmological scheme.

Two structural features deserve special notice.  First, the text employs
the \zh{七曜值日} (``Seven Luminaries Day-Rule''), the planetary week in
which each day and hour is governed in turn by one of the seven planets.
This is the Indian \textit{v\={a}ra} (planetary weekday) doctrine, which
reached China via the same Sogdian-Persian channel that transmitted the
\textit{Jiuzhi li} \citep{mak2024persian}.  The seven-day week is not
a native Chinese institution; its presence in the \zh{三辰通載} is a
direct inheritance from the Tang transmission, unmodified.

Second, the text contains the planetary period system known as
\zh{洞微百六限} (\textit{D\`ongw\=ei b\`aili\`u xi\`an},
``Penetrating-Subtlety Hundred-and-Six Limits'').  This is the Chinese
adaptation of the Indian \textit{da\'{s}\={a}} (time-lord) system, in which
each planet governs a defined span of a nativity's life in sequence.
The specific allocation of years and the sequencing of the seven governors
in the \zh{洞微百六限} shows closer affinity with the Indian
\textit{Yavanaj\={a}taka} tradition than with later purely Chinese
alternatives \citep{ho2003astrology}.

\subsection*{Identity Thread: The Four Remainders Crystallize}

The most consequential development in the \zh{三辰通載} for the purposes
of this paper is the crystallization of the \zh{四餘} as a canonical set.
By the Song period the four fictitious bodies have settled into their
classical configuration: \zh{羅睺} (Rahu, ascending node), \zh{計都}
(Ketu), \zh{月孛} (\textit{Yu\`eb\`ei}, ``Lunar Comet,'' mean apogee),
and \zh{紫氣} (\textit{Ziq\`{\i}}, ``Purple Vapour,'' a 28-year
intercalary point).

The identity of \zh{計都} in this period is significant.  As noted in
Section~\ref{sec:tang}, Niu Weixing's analysis of the 806~CE ephemeris
establishes that the apogee reidentification was already in place by
the end of the Tang.  The Song system inherits this identification
and embeds it in the stable four-body set.  The doubled apogee
tension---\zh{計都} as osculating apogee and \zh{月孛} as mean
apogee simultaneously present in the same system---is therefore a
structural feature of \zh{七政四餘} from its earliest codification,
not a later innovation.  We return to this tension in
Section~\ref{sec:yuan}.

\subsection*{Significance: A Baseline for Measuring Transformation}

The \zh{三辰通載} system is significant precisely as a \emph{baseline}.
It preserves the maximum Indian DNA that the Chinese tradition will
ever retain simultaneously in a single codified text.  Ecliptic
coordinates, sidereal zero-point, Jyotish-style house structure,
\textit{v\={a}ra} day-rule, \textit{da\'{s}\={a}}-derived period
system---all are present, with only modest adaptation.

This makes the \zh{三辰通載} methodologically invaluable.  The
Yuan equatorial reform (Section~\ref{sec:yuan}), the Ming quasi-ecliptic
synthesis (Section~\ref{sec:ming}), and the Jesuit tropical correction
(Section~\ref{sec:jesuit}) will successively overwrite exactly these
features.  Each subsequent transformation can be precisely characterized
by reference to what it departs from.  Without the Song baseline, the
later changes would be difficult to measure; with it, the direction and
magnitude of each transformation become legible.

In the language of the six-mode framework introduced in
Section~\ref{sec:intro}, the \zh{三辰通載} represents the terminal
state of Mode~1 (transliteration) and the opening condition for
Mode~3 (indigenous restructuring) that will follow in the Yuan.

% === §5 The Yuan Reform =======================================================
\section{Equatorial Sidereal: The Yuan Reform (\zh{鄭氏星案})}
\label{sec:yuan}
% ~1,500 words — Task 8

\subsection*{Transformation Mode~3: Indigenous Restructuring}

If Mode~1 (transliteration) is assimilation without digestion, and
Mode~2 establishes the baseline that transliteration produces, then
Mode~3 is the moment when the host culture begins actively reshaping
what it has received.  Indigenous restructuring does not reject the
imported content; it preserves the pseudo-planets, the period system,
the interpretive categories.  What it replaces is the \emph{form}---the
coordinate system in which that content is expressed.  The Yuan dynasty
(1271--1368~CE) provides the most dramatic instance of this mode in
the entire history of Chinese \zh{七政四餘}.

\subsection*{Historical Context: Guo Shoujing and the \zh{授時曆}}

The Yuan was an era of exceptional astronomical ambition.  Kublai
Khan's court sponsored simultaneous engagement with both the indigenous
Chinese tradition and the Islamic astronomical science arriving via
the westward connections of the Mongol empire.  An Islamic Astronomical
Bureau (\zh{回回司天監}, \textit{Huihu\'{i} S\={i}ti\={a}nj\={a}n})
operated in parallel with the traditional Chinese bureau, bringing
\textit{z\={\i}j} tables and armillary instruments of Persian and
Arabic origin into direct institutional contact with Chinese
practice \citep{mak2024persian}.

Within this environment, the astronomer Guo Shoujing (\zh{郭守敬},
1231--1316) led the most technically ambitious calendar reform in
Chinese history.  The resulting \zh{授時曆} (\textit{Sh\`oush\`{\i}
l\`{\i}}, ``Season-Granting Calendar,'' 1280) achieved solar-year
accuracy not surpassed in China for three centuries
\citep{sivin2009granting}.  Its technical foundation was the
equatorial coordinate system: Guo Shoujing designed a suite of
precision instruments, most famously the \zh{簡儀} (\textit{ji\v{a}n
y\'{i}}, ``simplified armillary''), that measured celestial positions
directly in right ascension and declination rather than converting
through the ecliptic.

The \zh{授時曆} also solved, for the first time in Chinese astronomy,
the mathematical problem of converting between ecliptic and equatorial
coordinates with precision.  Wagner's reconstruction of Guo Shoujing's
conversion algorithm demonstrates that the \zh{授時曆} employed a
spherical-trigonometric method---unprecedented in Chinese mathematical
astronomy---to map positions between the two frames
\citep{wagner2024gsj}.  Whether this mathematical capacity directly
influenced the \zh{七政四餘} practitioners who produced the
\zh{鄭氏星案} remains an open question; but the institutional and
intellectual environment in which equatorial measurement became the
dominant mode is the necessary precondition for the coordinate shift
we document.

\subsection*{The Text: \zh{鄭氏星案}}

The \zh{鄭氏星案} (\textit{Zh\`engsh\`{\i} x\={\i}ng'\`an},
``Zheng's Stellar Cases'') is a Yuan-dynasty \zh{七政四餘} manual
that represents the first systematic application of equatorial
coordinates to the inherited Indian astrological framework.\footnote{The
equatorial sidereal characterization of the \zh{鄭氏星案} awaits full
verification against the primary text; our characterization of its
coordinate system draws on secondary analyses, notably
\citealt{ho2003astrology} and \citealt{niu2010fourremains}.}
The title ``Stellar Cases'' suggests a practitioner's casebook of
worked examples---a format that carries the reader step by step through
astrological delineation---rather than a theoretical treatise.  This
makes it all the more striking that so fundamental a feature as the
coordinate system was reformulated.

\subsection*{Coordinate Thread: From Ecliptic to Equatorial}

The \zh{鄭氏星案} employs the \zh{赤道恆星製}
(\textit{ch\`{\i}d\`ao h\'{e}ngx\=ing zh\`{\i}},
``equatorial sidereal system'').  Where the Song \zh{三辰通載}
measured planetary longitudes along the ecliptic, the Yuan text
measures positions in right ascension, anchored to the native
Chinese system of twenty-eight lunar mansions (\zh{二十八宿},
\textit{\`ersh\'{i}b\={a} xi\`u}).  The twelve zodiacal houses,
in the Song system clean 30\textdegree{} ecliptic arcs,
become divisions of the celestial equator, irregular in ecliptic
terms but aligned with the observational infrastructure that Chinese
astronomy had built over millennia.

The conceptual consequences are far-reaching.  The ecliptic is the
plane of the zodiac; the zodiacal constellations that define Aries,
Taurus, Gemini, and the rest are ecliptic phenomena.  Once the
coordinate backbone is moved to the equator, the zodiacal constellation
concept does not merely become imprecise---it becomes inapplicable.
The \zh{鄭氏星案} does not attempt to preserve it.  What survives the
coordinate transformation is the functional infrastructure of
prediction: the seven governors, the four remainders, the period
system, the aspects.  What does not survive is the ecliptic framework
that gave the zodiacal houses their geometric meaning.

In the terminology of our six-mode framework, this is Mode~3 in its
purest form.  The \emph{content} of the imported system---pseudo-planets,
time-lords, interpretive techniques---persists intact.  The
\emph{form}---the coordinate operating system in which that content
runs---is replaced by the native Chinese alternative.  An analogy
from the history of computing is instructive: astronomical
``data'' has been successfully ported to a new ``operating system,''
but applications written natively for the old platform no longer
run without modification.  The zodiacal house system, native to the
ecliptic, cannot run on the equatorial platform without
fundamental redesign.

\subsection*{The Mechanism of Conversion}

How were positions translated from the old ecliptic sidereal frame
to the new equatorial sidereal frame in practice?  The \zh{授時曆}'s
conversion algorithm, reconstructed by Wagner, provides one plausible
pathway \citep{wagner2024gsj}.  Guo Shoujing's method works forward
from equatorial observations to ecliptic positions for calendar
purposes; the inverse problem---converting existing ecliptic-referenced
\zh{七政四餘} tables into equatorial coordinates---is mathematically
equivalent.

Whether \zh{七政四餘} practitioners had direct access to the
\zh{授時曆}'s computational machinery, or worked from
approximately-converted tables, or simply re-observed planetary
positions against the lunar mansion system, cannot be established
from the surviving text with certainty.  What can be established
is the outcome: the Yuan \zh{七政四餘} practitioner no longer
worked in ecliptic longitude but in a right-ascension equivalent
measured against the 28 lodges.  The practical effect was to make
\zh{七政四餘} positions directly comparable with, and perhaps
checkable against, the standard observational output of the
official astronomical bureau.

\subsection*{Identity Thread: The Doubled Apogee Persists}

The Yuan reform touches coordinates but not identity.  The \zh{四餘}
continue within the equatorial framework with their Song-period
identifications undisturbed.  \zh{羅睺} remains the ascending
node.  \zh{計都} remains the osculating lunar apogee.  \zh{月孛}
(\textit{Yu\`eb\`ei}) continues to track the mean apogee.
\zh{紫氣} (\textit{Ziq\`{\i}}) continues its 28-year cycle.

The doubled-apogee tension noted in Section~\ref{sec:song}---two
bodies, \zh{計都} and \zh{月孛}, both apsidal, one osculating and
one mean---therefore persists through the coordinate transformation.
This is not a logical contradiction within the system's own terms;
osculating and mean apogee are genuinely distinct quantities that
diverge by up to several degrees at any given moment.  But
from the perspective of an observer comparing the system with
Indian \textit{Jyotish}, the presence of two apogees where the
Indian system has one (the descending node) and one apogee
(\textit{mand\={o}cca}) requires explanation.  The Yuan text,
like the Song text before it, does not supply one.  The doubled
apogee is a structural feature that will be inherited, unchanged
in its identity if not in its coordinates, by the Ming synthesis.

\subsection*{Significance: The Most Dramatic Transformation}

Of the four coordinate regime changes traced in this paper, the
Yuan shift from ecliptic to equatorial is the most dramatic in
geometric terms.  The ecliptic and the equator intersect at an
angle of approximately 23.5\textdegree{} (the obliquity of the
ecliptic), which means that a position expressed in ecliptic
longitude and latitude cannot be converted to equatorial right
ascension and declination by a simple additive offset: the
conversion requires a rotation of the coordinate frame.  The
Song-to-Yuan transition therefore involved a genuine geometric
transformation of the entire positional database, not merely a
recalibration of a zero-point.

What makes this transformation analytically tractable is precisely
the stability of the content layer.  Because the pseudo-planets,
interpretive rules, and period allocations were preserved unchanged
across the coordinate shift, the transformation is detectable in
retrospect: one can compare the positions predicted by the Song
ecliptic model with those predicted by the Yuan equatorial model
for the same historical date, and the systematic offset will reveal
the nature of the conversion actually performed.  The computational
verification in Section~\ref{sec:computation} exploits exactly
this property.

The Yuan reform also opens a question that the text itself does
not answer: was the equatorial reframing driven by intellectual
preference---a desire to align \zh{七政四餘} with the prestige
of Guo Shoujing's new astronomical infrastructure---or by
practical necessity, as practitioners trained in the equatorial
tradition found ecliptic-referenced tables difficult to use?
The answer likely involves both, and the ambiguity is itself
historically significant.  What is clear is that after the Yuan
reform, the ecliptic sidereal frame of the Song baseline was
gone, replaced by a coordinate system whose deepest roots were
not Indian or Persian but natively Chinese.

% === §6 The Ming Synthesis ====================================================
\section{Quasi-Ecliptic Sidereal: The Ming Synthesis (\zh{果老星宗})}
\label{sec:ming}
% ~1,800 words — Task 9

\subsection*{Transformation Mode~4: Hybridization}

The Ming dynasty did not merely inherit the Yuan equatorial frame.  It
transformed it a second time, producing a coordinate system with no
counterpart in any other astronomical tradition---a system that looks
ecliptic-like on its surface but is derived from equatorial
measurements underneath.  This hybridization is the paper's most
important case study, and it is the juncture at which the
\zh{七政四餘} tradition achieves maximum distance from its Indian
original in both coordinate conventions and interpretive content.

\subsection*{The Canonical Texts}

The primary vehicle for this transformation is the
\zh{果老星宗} (\textit{Gu\v{o}l\v{a}o x\={\i}ngz\={o}ng}, ``Star-School
of the Elder of the Fruit''), the Ming-dynasty canonical text of the
\zh{七政四餘} tradition.\footnote{The \zh{果老星宗} takes its name
from Zhang Guolao (\zh{張果老}), one of the Eight Immortals of Daoist
legend and a figure credited, in mythologised form, with transmitting
astral knowledge from antiquity.  The attribution is hagiographic
rather than historical.}  The text compiles computational rules,
interpretive glosses, and worked examples for all eleven bodies of
the mature system.  It is the text that most modern Chinese
astrologers regard as the authoritative statement of ``standard''
\zh{七政四餘} practice, and its coordinate conventions are therefore
the ones most consequential for the living tradition.

A near-contemporary source of comparable authority is the
\zh{星學大成} (\textit{X\={\i}ngxu\'e d\`ach\'eng}, ``Great Compendium
of Astral Studies'') by Wan Minying \zh{萬民英} (1521--1603)
\cite{ho2003astrology}.  Wan's compilation is more systematic in its
presentation of computational methods, and the two texts are
frequently cited together as the twin pillars of Ming
\zh{七政四餘} scholarship.  Where they agree---and on the coordinate
conventions they agree---we can speak of a Ming consensus.

\subsection*{Coordinate Thread: Defining the \zh{似黃道恆星製}}

The coordinate system of the Ming texts is the
\zh{似黃道恆星製} (\textit{s\`{\i} hu\'angd\`ao h\'engx\={\i}ng
zh\`{\i}}, ``quasi-ecliptic sidereal system'').  The name encodes
both an assertion and a qualification: the system is \textit{like}
(\zh{似}) the ecliptic---it maps positions onto zodiac-style signs
and houses---but the underlying measurements are taken along the
equator, not the ecliptic.  No other astronomical tradition has
produced such a hybrid.

\subsubsection*{Mathematical Definition}

The construction proceeds in three steps.  First, the equator is
divided into twelve equal houses of $30\degree$ each, beginning at
the vernal equinox.  Second, each equatorial boundary is projected
onto the ecliptic plane using the standard transformation between
right ascension $\alpha$ and ecliptic longitude $\lambda$ at zero
ecliptic latitude ($\beta = 0$):
\begin{equation}
  \tan\alpha = \cos\varepsilon \cdot \tan\lambda
  \label{eq:ra-to-ecl}
\end{equation}
where $\varepsilon \approx 23.5\degree$ is the obliquity of the
ecliptic.  Inverting to obtain the ecliptic boundary corresponding to
each equatorial house boundary:
\begin{equation}
  \lambda = \arctan\!\left(\frac{\sin\alpha}{\cos\alpha \cdot \cos\varepsilon}\right)
  \label{eq:ecl-boundary}
\end{equation}
Third, planetary positions, originally computed in equatorial
coordinates, are re-expressed in terms of whichever ecliptic house
they fall into under the projected boundaries.

The result is that the twelve ``signs'' look ecliptic---they are
labelled with the same zodiacal names as in any ecliptic
system---but their boundaries are not equally spaced in ecliptic
longitude.  The unequal spacing is a direct consequence of the
obliquity $\varepsilon$.  Houses near the equinoxes, where the
equator and ecliptic are most nearly parallel in their rate of
longitude change, acquire wider ecliptic extents; houses near the
solstices, where the relationship is most compressed, acquire
narrower extents.

\subsubsection*{Computed Boundaries and the Symmetry Pattern}

Table~\ref{tab:quasi-ecliptic} gives the computed ecliptic boundaries
for all twelve houses under obliquity $\varepsilon = 23.5\degree$
(appropriate for approximately 1400~CE).

\begin{table}[ht]
\centering
\caption{Quasi-ecliptic (\zh{似黃道}) house boundaries compared to equal ecliptic divisions. Obliquity $\varepsilon = 23.5\degree$ (circa 1400~CE). Houses near equinoxes are wider ($\sim$32.2\degree); houses near solstices are narrower ($\sim$27.9\degree).}
\label{tab:quasi-ecliptic}
\small
\begin{tabular}{@{}rrrr@{}}
\toprule
House & RA boundary & Ecliptic boundary & Width \\
\midrule
1 & 0.0\degree & 0.00\degree & 32.19\degree \\
2 & 30.0\degree & 32.19\degree & 29.91\degree \\
3 & 60.0\degree & 62.10\degree & 27.90\degree \\
4 & 90.0\degree & 90.00\degree & 27.90\degree \\
5 & 120.0\degree & 117.90\degree & 29.91\degree \\
6 & 150.0\degree & 147.81\degree & 32.19\degree \\
7 & 180.0\degree & 180.00\degree & 32.19\degree \\
8 & 210.0\degree & 212.19\degree & 29.91\degree \\
9 & 240.0\degree & 242.10\degree & 27.90\degree \\
10 & 270.0\degree & 270.00\degree & 27.90\degree \\
11 & 300.0\degree & 297.90\degree & 29.91\degree \\
12 & 330.0\degree & 327.81\degree & 32.19\degree \\
\bottomrule
\end{tabular}
\end{table}

The symmetry pattern is instructive.  Houses~1, 6, 7, and~12 (near
the equinoxes) are the widest at $\approx 32.19\degree$; houses~3,
4, 9, and~10 (near the solstices) are the narrowest at
$\approx 27.90\degree$.  The intermediate houses (2, 5, 8, 11) fall
at $\approx 29.91\degree$.  The pattern has period-two symmetry
because the obliquity effect is symmetric about both equinox lines:
the ecliptic and equator cross at the same angle at Aries~$0\degree$
and at Libra~$0\degree$, and the same compression applies at the two
solsticial points.

The maximum deviation from the equal $30\degree$ division is
$\approx \pm 2.19\degree$.  This seems small.  It is not
astronomically negligible.  A body at the boundary between two houses
can be placed in either house depending solely on which convention is
applied.  For slow-moving bodies such as Saturn (average motion
$\approx 0.033\degree$/day), the $\pm 2.19\degree$ discrepancy
corresponds to a positional ambiguity of roughly 66 days.  For the
nodes (average regression $\approx 0.053\degree$/day) the ambiguity
is approximately 41 days.  The quasi-ecliptic frame therefore
introduces a systematic, predictable source of inter-system
discrepancy that is distinct from---and additive with---the
coordinate-zero-point discrepancies discussed in
Sections~\ref{sec:song} and~\ref{sec:jesuit}.

\subsubsection*{Why the System Is Novel}

It is worth pausing on this point.  Indian astrology uses ecliptic
coordinates directly.  Greek astrology uses ecliptic coordinates
directly.  Persian and Arabic astrology---the traditions that served
as intermediaries for the Tang transmission---use ecliptic
coordinates directly.  The Yuan reform replaced ecliptic with
equatorial coordinates.  The Ming synthesis did something none of
these did: it used equatorial measurements to generate apparently
ecliptic house divisions.  The \zh{似黃道恆星製} is a genuinely new
coordinate species, not a corruption or a simplification of any
predecessor system.

\subsection*{Identity Thread: The Mature \zh{四餘} System}

The Ming texts present the most fully elaborated version of the
four-remainder (\zh{四餘}) system.  The four bodies are now formally
defined, their computational methods codified, and their astrological
significances established.  The two bodies whose identities were
contested in earlier periods now coexist in a stable, if
theoretically uneasy, configuration.

\zh{羅睺} (\textit{R\'ahu}) retains the identity it has held since
the Tang transmission: the ascending lunar node.  No reidentification
touches it.  It remains the closest of the four bodies to its Indian
original in both name and astronomical definition.

\zh{計都} (\textit{K\`et\`u}) is the osculating lunar apogee---the
reidentification already in place by the end of the Tang and
confirmed by the Song codification.  In the Ming texts this identity
is simply assumed, not argued; the apogee assignment has become the
tradition.

\subsubsection*{The Mean Apogee: \zh{月孛}}

\zh{月孛} (\textit{Yu\`eb\`o\`{\i}}, ``lunar comet'' or ``lunar
apoapsis'') is the mean lunar apogee---the apogee averaged over the
Moon's short-period perturbations and free of the oscillations that
distinguish the osculating apogee.  Its identification with Zhao
Youqin (\zh{趙友欽}, fl.\ late 13th--early 14th century) represents
one of the genuinely indigenous contributions to the system
\cite{niu2010fourremains}.  Zhao's derivation proceeded from careful
observation of the Moon's varying daily motion: slower motion near
apogee, faster near perigee.  By tracking the Moon's speed across
many lunations, the mean apogee longitude can be extracted without
requiring explicit knowledge of the underlying orbital geometry.
The result is a quantity distinct from \zh{計都} precisely because
it is mean rather than osculating.

\subsubsection*{The Purple Qi: \zh{紫氣}}

\zh{紫氣} (\textit{Z\v{\i}q\`{\i}}, ``Purple Qi'') is a fictitious
point with a cycle of approximately 28 years.  It has no
astronomical counterpart---it does not correspond to any orbital
element of the Moon, the Sun, or any planet---and it has no Indian
precedent.  It is a purely Chinese invention, its cycle apparently
chosen to be commensurate with Jupiter's 12-year cycle and Saturn's
29.5-year cycle in a numerologically satisfying way.  Unlike the
other three remainders, \zh{紫氣} cannot be confirmed or disconfirmed
by independent astronomical observation; its position is determined
entirely by the calendrical formula that defines it.

\subsubsection*{The Doubled Apogee Problem}

The Ming system thus contains two bodies that both track the lunar
apse line: \zh{計都} (osculating apogee) and \zh{月孛} (mean
apogee).  They inhabit the same orbital degree of freedom.  At any
given moment they differ by a quantity that is the difference between
the osculating and mean lunar apogee---a difference that varies over
the anomalistic month and reaches a maximum of several degrees.

This structural redundancy has never been satisfactorily resolved
within the tradition.  No Ming or subsequent text provides a rigorous
explanation of why both bodies are necessary, or what astrological
distinction should be drawn between them.  The doubled apogee
remains an open theoretical question \cite{niu2010fourremains}.

\subsubsection*{The Beginning of Bazi-ification}

One further development of the Ming period deserves mention, even
though its full consequences belong to a later section.  Mak (2014)
documents the emergence of a practice he terms ``bazi-ification'':
the derivation of \zh{四餘} positions from \zh{八字}
(\textit{b\={a}z\`{\i}}, ``eight characters'' or ``four pillars'')
numerology rather than from astronomical ephemerides
\cite{mak2014pseudoplanets}.  In this approach, a practitioner
determines the positions of the four remainders not by computing
their actual positions in the sky but by applying formulas that take
the birth date and time as input and yield positions by arithmetic
rule.  The positions so obtained are not guaranteed to match the
astronomical bodies they nominally represent.

This is the first step toward the algorithmic fossilization that
will be examined in Section~\ref{sec:fossil}.  At the Ming stage it
is a minority practice competing with genuine ephemeris-based
computation.  Its significance is not yet apparent.

\subsection*{Significance: High-Water Mark of Sinicization}

The Ming synthesis represents the maximum distance the
\zh{七政四餘} system ever achieved from its Indian original,
measured along both the coordinate and identity threads simultaneously.

On the coordinate thread, the \zh{似黃道恆星製} is the culmination
of a three-stage transformation.  The Song system preserved the
Indian ecliptic frame.  The Yuan reform replaced it with the native
Chinese equatorial frame.  The Ming synthesis performed a further
operation: it used the equatorial frame to construct an
\textit{apparent} ecliptic frame whose divisions are unequal in
ecliptic longitude.  The result superficially resembles the ecliptic
systems of the Indian, Greek, and Arabic traditions but is not
reducible to any of them.  Mode~4 (hybridization) is the appropriate
designation: the imported ecliptic appearance and the native
equatorial foundation are fused into something genuinely new.

On the identity thread, all four remainders are now present, two of
them (\zh{月孛} and \zh{紫氣}) without Indian precedent.  The
tradition has added bodies from native Chinese sources and retained
them in permanent parallel with the bodies inherited from India.

Yet the system has not yet fossilized.  The \zh{果老星宗} and
\zh{星學大成} both give computational procedures that, however
hybrid, are in principle checkable against the sky.  Positions are
derived from genuine orbital elements, even if the coordinate frame
in which they are expressed is novel.  The astronomical content
survives.  The next disruption will come not from Chinese synthesis
but from European imposition \cite{mak2014pseudoplanets,niu2010fourremains}.

% === §7 The Jesuit Disruption =================================================
\section{Ecliptic Tropical: The Jesuit Disruption (1645)}
\label{sec:jesuit}
% ~2,000 words — Task 9

\subsection*{Transformation Mode~5: Foreign Correction}

In 1645 Johann Adam Schall von Bell (\zh{湯若望}), a German Jesuit
astronomer serving as director of the imperial Bureau of Astronomy,
promulgated a reformed astronomical system under the new Qing dynasty.
The reform was the most violent single disruption in the history of
the \zh{七政四餘} tradition---more disruptive, in the short term,
than even the Tang transmission---because it changed both the
coordinate thread and the identity thread simultaneously, imposed
from outside the tradition by institutional authority rather than
emerging from internal Chinese scholarly development.

This is Mode~5: foreign correction.  Unlike Mode~3 (indigenous
restructuring, which was organic cultural pressure adapting an
imported system to native categories), Mode~5 is a deliberate,
externally authorized reset.  The new system was not proposed for
scholarly debate; it was mandated by imperial decree.

\subsection*{The Reform and Its Context}

The institutional context matters.  The late Ming dynasty had
suffered a series of calendar prediction failures that undermined
the credibility of the existing astronomical administration.  When
the Qing forces captured Beijing in 1644, Schall---who had been in
China since 1622 and had already demonstrated the accuracy of
European methods in predicting eclipses---was positioned to offer an
alternative.  In 1645 the Shunzhi Emperor promulgated the
\zh{時憲曆} (\textit{Sh\`{\i}xi\`an l\`{\i}}, ``Calendar of Edicts
Conforming to Heaven''), the new official calendar based on European
methods \cite{liu2020missionary}.

The computational infrastructure underlying the calendar was drawn
from the Tychonic system: Tycho Brahe's geo-heliocentric model,
which placed the planets in orbit around the Sun while maintaining
the Sun in orbit around the Earth.  From a positional astronomy
standpoint, the Tychonic system is mathematically equivalent to the
Copernican heliocentric model for the purposes of computing planetary
longitudes, and its predictions were substantially more accurate
than those of the Chinese \zh{授時曆} over the century of positional
drift that had accumulated since the Yuan reform \cite{jami2015calendar}.

For the \zh{七政四餘} system specifically, the reform imposed a
coordinate change of fundamental character: from sidereal to tropical.
The zero point of longitude shifted from fixed stars to the vernal
equinox.  The new system can be designated the
\zh{黃道迴歸製} (\textit{hu\'angd\`ao hu\'{\i}gu\={\i} zh\`{\i}},
``ecliptic tropical system'').

\subsection*{Coordinate Thread: The Ironic Completion}

The coordinate history of \zh{七政四餘} traces a closed trajectory
whose ironic structure is only visible from the vantage point of
the Qing reform.

The Song system used \textit{ecliptic sidereal} coordinates: the
ecliptic plane, zero point at a fixed star.  The Yuan reform
shifted to \textit{equatorial sidereal} coordinates: the equatorial
plane, zero point still at a fixed star.  The Ming synthesis
produced the \textit{quasi-ecliptic sidereal} hybrid: equatorial
measurements mapped to ecliptic-like divisions, zero point at a
fixed star.  The Qing reform imposed \textit{ecliptic tropical}
coordinates: the ecliptic plane, zero point at the vernal equinox.

The trajectory thus moves from ecliptic sidereal (Song) to equatorial
sidereal (Yuan) to quasi-ecliptic sidereal (Ming) and then snaps
back to the ecliptic plane (Qing)---but with the zero point
replaced.  The system ended in the ``opposite corner'' from where
it began: same plane as the Indian original (ecliptic), same plane
as the Song codification, but a different zero point.

The magnitude of the coordinate zero-point discrepancy is
substantial.  The tropical-sidereal difference is the precession
accumulated since the epoch at which the sidereal zero was defined.
By 1645, this amounts to approximately $24\degree$.  A planet at
\textit{tropical} $\lambda = 30\degree$ (the start of Taurus in
the tropical frame) is located at approximately \textit{sidereal}
$\lambda = 6\degree$ (still in Aries in the sidereal frame).  All
eleven bodies of the \zh{七政四餘} system shift by roughly one
zodiac sign under the change of convention.

This $\approx 24\degree$ discrepancy is the coordinate source of
the inter-system disagreements documented in the companion paper
\cite{liu2020missionary}.  When modern practitioners inadvertently
mix the Ming (sidereal) and Qing (tropical) conventions, a
systematic error of this magnitude is introduced into every
positional judgment.

\subsection*{Identity Thread: Ketu Restored, \zh{紫氣} Abolished}

The Jesuit reform was no less consequential on the identity thread.
Two changes stand out.

First, \zh{計都} reverted to the descending lunar node.  The
apogee reidentification that had persisted from the Tang period
through the Song and Ming codifications---a reidentification
demonstrated by Niu Weixing's computational analysis
\cite{niu1995ketu}---was reversed.  In the Schall reform,
\zh{計都} is once again the descending node, its original Indian
identity restored after approximately 900 years in the guise of
the apogee.  The reversal was not the result of Chinese scholarly
debate about the original Indian sources; it was a consequence of
the imported European computational framework, which had no concept
of an ``apogee body'' and naturally assigned the node bodies their
standard Indian identities.

Second, \zh{紫氣} was abolished entirely.  The purely Chinese
fictitious point with its 28-year cycle had no place in the
European-derived framework.  Its removal was a decisive reduction of
the tradition's indigenous Chinese additions, eliminating the one
body that had been added by Chinese astronomers with no foreign
precedent.

An elemental inconsistency followed from the \zh{計都} reidentification.
Within the \zh{五行} (\textit{w\v{u}x\'{\i}ng}, ``five phases'')
assignments that govern astrological interpretation, the Cai Boli
(\zh{蔡伯勵}) tradition---which preserves the Qing convention and
remains influential in Cantonese-speaking communities---encodes a
fire/earth swap relative to the earlier convention.  Because
\zh{計都} as apogee and \zh{計都} as descending node occupy
different positions in the sky at the same moment, their
phase-element affinities, derived from their positions in the
annual cycle, differ systematically.  The resulting
\zh{計北羅南}/\zh{計南羅北} toggle (``Ketu north, Rahu south'' vs.\
``Ketu south, Rahu north'') that appears in modern software
encoding reflects not a computational error but a genuine dual
inheritance from the pre- and post-Jesuit conventions
\cite{liu2020missionary}.

\subsection*{The Calendar Case (1664--1669)}

The Jesuit reform did not go unchallenged.  In 1664, Yang Guangxian
(\zh{楊光先}), a Han Chinese official and amateur astronomer,
submitted a memorial to the Kangxi Emperor accusing the Jesuits
of errors in their astronomical methods, of promoting heterodox
cosmology (the heliocentric model was philosophically suspect in
the Confucian framework), and of using their calendar monopoly to
subvert Chinese institutions \cite{jami2015calendar}.

The accusation had political as well as astronomical dimensions.
The Kangxi Emperor was a child; the regency was dominated by
conservative bannermen suspicious of European influence.  In 1665
Schall was convicted, sentenced to death, and subjected to the
\textit{lingchi} (``death by a thousand cuts'').  The sentence was
commuted---according to contemporary accounts, because a severe
earthquake struck Beijing shortly after sentencing, interpreted as
a sign of heaven's displeasure at the verdict.  Schall died under
house arrest later that year.  Yang Guangxian was appointed to lead
the Bureau of Astronomy, reverting briefly to Chinese computational
methods.

The scientific resolution came in 1669.  The Kangxi Emperor, now
governing in his own right and genuinely curious about mathematics
and astronomy, arranged a public astronomical competition.  Ferdinand
Verbiest (\zh{南懷仁}), a Flemish Jesuit who had served under
Schall, challenged Yang Guangxian to predict the length of a shadow
cast by a gnomon at noon on a specified date, the position of the
Sun, Moon, and visible planets for a specified day, and the timing
of an impending lunar eclipse.  Verbiest's predictions, based on
European methods, were consistently more accurate.  Yang Guangxian
conceded, resigned his post, and was escorted home in disgrace
\cite{jami2015calendar}.

Verbiest was appointed to lead the Bureau of Astronomy.  The
European astronomical infrastructure was reinstated.  The Qing
\zh{七政四餘} convention---ecliptic tropical coordinates, Ketu as
descending node, no \zh{紫氣}---was confirmed as the official
standard.

\subsection*{The Kangxi Synthesis and Pragmatic Dualism}

The Kangxi Emperor's resolution of the Calendar Case was not merely
a vindication of European methods.  It was the articulation of a
pragmatic dualism that persists in the tradition to this day: Western
astronomical data for computational precision, Chinese astrological
framework for interpretive content.

The Emperor did not mandate that Chinese practitioners adopt European
cosmology.  He mandated that official calendrical computation use
European positional astronomy.  The astrological interpretations
attached to planetary positions---the \zh{五行} element assignments,
the house-by-house delineations, the fate-technical categories
inherited from the medieval Chinese synthesis---remained Chinese.
The result was a hybrid of a different kind from the Ming
\zh{似黃道恆星製}: not a mathematical hybrid of coordinate systems
but an institutional hybrid of computational authority and
interpretive sovereignty \cite{liu2020missionary,jami2015calendar}.

\subsection*{Significance: The Ironic Circle}

The Jesuit disruption completes what may be called the coordinate
irony of \zh{七政四餘} history.  The tradition began with ecliptic
coordinates imported from India.  Over six centuries it migrated
away from the ecliptic, first to the equator (Yuan), then to the
quasi-ecliptic hybrid (Ming), achieving maximum sinicization.  The
Jesuit reform pulled it back to the ecliptic plane---but not to the
Indian ecliptic.  The new ecliptic system had a European zero point
(vernal equinox) rather than a sidereal Indian zero point, making
it tropical rather than sidereal.  The coordinate trajectory
returned to its plane of origin but not to its origin.

The contrast with Mode~3 (indigenous restructuring) illuminates what
makes Mode~5 distinctive.  The Yuan reform was driven by Chinese
astronomers applying the tools of Chinese observational astronomy to
an inherited system; the result was organic, internally motivated,
and absorbed over several generations.  The Qing reform was driven
by Jesuit astronomers with institutional backing from the imperial
court; the result was abrupt, externally imposed, and generated a
fractured dual tradition that survives into the present.

The fracture is visible in modern practice.  The Cai Boli tradition
(predominantly Cantonese, post-Qing) uses tropical coordinates and
identifies \zh{計都} as the descending node.  The mainstream
\zh{果老星宗} tradition (drawing on Ming texts) uses sidereal
coordinates and identifies \zh{計都} as the apogee.  Both claim
historical legitimacy.  Both are historically legitimate.  The
$\approx 24\degree$ positional discrepancy between them is not an
error; it is the fossil record of 1645.

% === §8 Algorithmic Fossilization =============================================
\section{Algorithmic Fossilization}\label{sec:fossil}
% ~1,200 words

\subsection*{Mode~6: From Living Practice to Frozen Algorithm}

The sixth and final transformation mode differs fundamentally from those that
preceded it.  Transliteration, reidentification, indigenous restructuring,
hybridization, and foreign correction were all active interventions: scholars
consciously adapted inherited material to new intellectual frameworks.
Algorithmic fossilization is a passive process---astronomical content is not
transformed but \emph{forgotten}, while its terminological shell persists
unchanged.  The result is a system that presents an astronomical facade over an
empty interior.

\subsection*{Bazi-ification of the Seven Governors}

The process began quietly during the Ming dynasty.  As \zh{七政四餘} practice
spread beyond the circle of court astronomers who had access to working
ephemerides, a numerological shortcut emerged: deriving planetary positions not
from actual astronomical computation but from the \zh{八字}
(``eight characters''---the four pairs of heavenly stem and earthly branch that
encode the birth date and time in the sexagenary cycle).

Mak~\cite{mak2014pseudoplanets} documents this process in detail.  In the
authentic tradition, a \zh{七政四餘} chart requires computing the ecliptic
longitudes of seven planets and four pseudo-planets at the moment of birth
using ephemeris tables---a procedure whose complexity rivalled anything in
contemporary European astrology.  In the degraded tradition, the positions are
``looked up'' from tables indexed by the \zh{八字} itself, making the result a
pure function of the birth year-month-day-hour encoding rather than of the
physical sky.

The astronomical content is progressively displaced by its numerological
surrogate at three stages.  First, practitioners consult pre-computed position
tables arranged by year-branch rather than by Julian date.  Second, the tables
themselves are generated not from planetary theory but by extrapolating mean
motions quantized to sexagenary intervals.  Third---and most damaging---the
pseudo-planets \zh{羅睺} (Rahu) and \zh{計都} (Ketu) acquire fixed angular
separations in these tables, destroying the independent variation that is their
only diagnostic value.  Mak calls this the ``\zh{八字} colonization'' of the
\zh{七政四餘} system: the birth chart format survives while its astronomical
substance is replaced by recycled horoscope numerology.

This transformation is the most insidious of the six modes precisely because it
is invisible from the outside.  A fossilized chart displays the same labels, the
same twelve-house wheel, and the same seven-plus-four bodies as an
astronomically correct chart.  Only a comparison against independent ephemeris
data reveals the substitution.

\subsection*{Modern Software and the Contested Toggle}

The fossilization process reached its contemporary endpoint in the wave of
Chinese astrology software that emerged in the 1990s and 2000s.  Modern
implementations---including the widely used
\textsc{MOIRA} suite (\texttt{BahnAstro/MOIRA\_chinese\_astrology} on
GitHub)---inherit the Qing-period Jesuit convention and define \zh{計都} as the
\emph{descending} lunar node.  This was, as documented in Section~\ref{sec:qing},
the explicit policy of Schall von Bell's \textit{Shixian} calendar of 1645, and
it contradicts the pre-Qing scholarly consensus recovered by Niu
Weixing~\cite{niu1995ketu}.

A revealing symptom of the underlying confusion is the
\zh{計北羅南}/\zh{計南羅北} toggle that appears in several software packages.
In Chinese this toggle reads ``Ketu-north Rahu-south'' versus ``Ketu-south
Rahu-north,'' and users are invited to choose between them as a matter of
preference or school affiliation.  The toggle exists because practitioners have
collectively sensed that the node assignment is contested across lineages, but
the astronomical literacy required to adjudicate the question---which would
require comparing chart positions against computed ephemerides---has been lost.
The toggle thus freezes a historically produced ambiguity into a user-interface
affordance, institutionalizing ignorance as optionality.

A second pathology concerns coordinate mixing.  Several implementations compute
planetary longitudes using the \zh{似黃道} (quasi-ecliptic sidereal) system
inherited from the Ming \zh{果老星宗} tradition, while computing the solar and
lunar positions in tropical coordinates---a difference of approximately~$24\degree$
at the present epoch.  Because neither developer nor user retains the conceptual
vocabulary to distinguish sidereal from tropical frames, the discrepancy goes
undetected.  The~$24\degree$ ayanamsa offset, which represents the accumulated
precession since the Spica-anchored epoch of the \textit{siddh\=anta} tradition,
is treated as a bug report rather than recognized as the signature of a
coherent---if now misapplied---reference frame.

\subsection*{The Archaeological Record in Code}

Mode~6 completes the transformation cycle.  The astronomical substance that
survived transliteration from Sanskrit, reidentification from node to apogee,
indigenous restructuring into the twelve-branch system, hybridization with the
quasi-ecliptic coordinate, and Jesuit correction back to the node is finally
lost---not through active intervention but through passive forgetting across
generations of practitioners who transmitted the form without the content.

Yet the very layers of misunderstanding preserved in modern software constitute
an unintentional archive.  The \zh{計北羅南} toggle records the memory of a
contested scholarly tradition.  The mixed-coordinate pathology records the
seam between the Ming and Qing epochs.  The \zh{八字}-derived position tables
record the social expansion of the practice beyond its astronomical
infrastructure.  Like geological strata, each layer of the software encodes the
conditions under which it was deposited.  The task of historical reconstruction
is to read those strata in reverse---not to condemn the modern practitioner, but
to recover the original astronomical content from its fossilized
residue~\cite{mak2014pseudoplanets}.

% === §9 Computational Verification ============================================
\section{Computational Verification}\label{sec:computation}
% ~1,500 words

\subsection*{Verification~1: Ketu Period Analysis}

The core claim of this paper---that the \zh{計都} of Tang-dynasty ephemerides
tracks the osculating lunar apogee rather than the descending node---rests
ultimately on numerical evidence.  We reproduce Niu Weixing's key finding using
Swiss Ephemeris algorithms (Meeus polynomial series for the lunar perigee and
node) and extend it over the full 62-year span of the \zh{七曜攘災訣} ephemeris
(806--868~CE).

The two candidate bodies have fundamentally different periods.  The osculating
lunar apogee (\textit{apogee convention}) completes one full revolution of the
ecliptic in $\approx 8.85$~years; the descending node (\textit{node convention})
completes one revolution in $\approx 18.6$~years (retrograde).  Over the
62-year span of the Tang ephemeris, the apogee-convention Ketu therefore
completes $\approx 7.0$ full revolutions while the node-convention Ketu
completes only $\approx 3.3$.  These are not approximately equal trajectories
with a constant offset: they are \emph{incommensurable} within any astronomical
epoch shorter than their least common multiple ($\approx 52$~years).

Figure~\ref{fig:ketu-period} plots the ecliptic longitudes of both bodies from
806 to 816~CE at 120-day intervals.  The apogee trace advances rapidly and
completes nearly one full revolution in the plotted decade; the node trace
retreats uniformly and covers only $\approx 200\degree$ in the same interval.
The angular divergence between the two conventions exceeds $90\degree$ for
large portions of this window.  Any historical ephemeris whose tabulated
positions match one trajectory \emph{cannot} match the other: the two
hypotheses are empirically distinguishable with positional precision far
below the accuracy of medieval instruments.

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=\columnwidth,
    height=6cm,
    xlabel={Year (CE)},
    ylabel={Ecliptic Longitude ($\degree$)},
    xmin=806, xmax=816.5,
    ymin=0, ymax=380,
    ytick={0,60,120,180,240,300,360},
    minor ytick={30,90,150,210,270,330},
    grid=both,
    grid style={line width=0.3pt, draw=gray!30},
    major grid style={line width=0.5pt, draw=gray!50},
    legend pos=north west,
    legend style={font=\small},
    tick label style={font=\small},
    label style={font=\small},
]
% Apogee data (every 120 days, 806--816 CE)
\addplot[color=apogeeblue, thick, mark=none] coordinates {
    (806.00, 279.2) (806.33, 293.3) (806.66, 304.1) (806.99, 320.5)
    (807.31, 331.1) (807.64, 346.2) (807.97, 359.2) (808.30,  12.0)
    (808.63,  27.0) (808.96,  37.7) (809.29,  54.1) (809.61,  65.1)
    (809.94,  79.1) (810.27,  93.5) (810.60, 104.6) (810.93, 119.2)
    (811.25, 131.3) (811.58, 140.5) (811.91, 155.8) (812.24, 168.0)
    (812.57, 176.2) (812.89, 190.7) (813.22, 203.8) (813.55, 211.6)
    (813.88, 224.4) (814.20, 237.7) (814.53, 245.1) (814.86, 256.0)
    (815.19, 269.9) (815.52, 278.3) (815.84, 287.1) (816.17, 300.1)
};
\addlegendentry{Apogee convention ($T \approx 8.85$~yr)}
% Node data (every 120 days, 806--816 CE)
\addplot[color=nodered, thick, dashed, mark=none] coordinates {
    (806.00,  80.8) (806.33,  74.5) (806.66,  68.1) (806.99,  61.7)
    (807.31,  55.4) (807.64,  49.0) (807.97,  42.8) (808.30,  36.4)
    (808.63,  30.1) (808.96,  23.8) (809.29,  17.4) (809.61,  11.1)
    (809.94,   4.7) (810.27, 358.4) (810.60, 352.0) (810.93, 345.7)
    (811.25, 339.3) (811.58, 332.9) (811.91, 326.6) (812.24, 320.2)
    (812.57, 313.8) (812.89, 307.5) (813.22, 301.1) (813.55, 294.8)
    (813.88, 288.4) (814.20, 282.0) (814.53, 275.7) (814.86, 269.3)
    (815.19, 263.0) (815.52, 256.6) (815.84, 250.3) (816.17, 243.9)
};
\addlegendentry{Node convention ($T \approx 18.6$~yr, retrograde)}
\end{axis}
\end{tikzpicture}
\caption{Ecliptic longitudes of the two candidate \zh{計都} bodies, 806--816~CE,
sampled every 120~days.  The apogee-convention body (solid blue) advances
prograde and completes nearly one full revolution; the node-convention body
(dashed red) retreats retrograde, covering $\approx 200\degree$ in the same
interval.  The two trajectories diverge by more than $90\degree$ for most of
the plotted window, making them empirically distinguishable from historical
ephemeris data.  Data generated from Meeus polynomial series via Swiss
Ephemeris; \textit{cf.}~\cite{niu1995ketu}.}
\label{fig:ketu-period}
\end{figure}

\subsection*{Verification~2: Four-Regime Coordinate Comparison}

To quantify the practical consequences of coordinate-system choice for chart
interpretation, Table~\ref{tab:four-regime} presents the computed positions of
the seven standard \zh{七政四餘} bodies for a reference date of 1400-06-15
(Beijing, noon local time) in all four coordinate regimes documented in
this paper.

\begin{table*}[ht]
\centering
\caption{Ecliptic and equatorial positions of the \zh{七政四餘} bodies for
1400-06-15 (Beijing, noon) in four coordinate regimes.  ``Ecl.\ Sid.'' = ecliptic
sidereal (\zh{三辰通載} / Song regime); ``Eq.\ Sid.'' = equatorial sidereal
(\zh{鄭氏星案} / Yuan regime); ``Quasi-ecl.'' = \zh{似黃道} sidereal
(\zh{果老星宗} / Ming regime, with house shown); ``Tropical'' = ecliptic tropical
(\textit{Shixian} / Qing--Jesuit regime).  All values in degrees.}
\label{tab:four-regime}
\small
\begin{tabular}{lrrrr}
\toprule
\textbf{Body} & \textbf{Ecl.\ Sid.} & \textbf{Eq.\ Sid.} & \textbf{Quasi-ecl.\ Sid.} & \textbf{Tropical} \\
\midrule
Sun          & 250.6 & 250.0 & 250.6 (H8)  &  83.5 \\
Moon         &  27.4 &  24.9 &  27.4 (H1)  & 220.3 \\
Mars         & 117.8 & 120.2 & 117.8 (H4)  & 310.7 \\
Jupiter      & 212.3 & 209.8 & 212.3 (H7)  &  45.2 \\
Rahu         & 327.1 & 328.6 & 327.1 (H11) & 160.0 \\
Ketu (apogee)& 259.5 & 259.7 & 259.5 (H9)  &  92.4 \\
Ketu (node)  & 147.1 & 148.6 & 147.1 (H5)  & 340.0 \\
\bottomrule
\end{tabular}
\end{table*}

Several features of Table~\ref{tab:four-regime} deserve comment.  First, the
ecliptic sidereal and equatorial sidereal regimes differ by up to $2.5\degree$
(Mars: $117.8\degree$ vs.\ $120.2\degree$), a gap that is small enough to be
invisible to pre-modern observational methods yet large enough, at Saturn's mean
motion of $\approx 0.033\degree$/day, to shift a slowly moving planet across a
house boundary over a span of several weeks.  The Yuan shift from ecliptic to
equatorial measurement was therefore not an innocuous change of notation but a
physically meaningful displacement of planetary positions within the chart.

Second, the tropical column---representing the Jesuit \textit{Shixian} regime---is
offset from all three sidereal columns by $\approx 167\degree$, the approximate
complement of the~$\sim 24\degree$ ayanamsa at this epoch.  A practitioner
comparing a Ming sidereal chart against a Qing tropical chart without awareness
of this offset would conclude that every planet had been displaced by nearly half
the zodiac.

Third, and most diagnostically: the two Ketu entries differ by $112.4\degree$
in the ecliptic sidereal regime ($259.5\degree$ vs.\ $147.1\degree$) and by
$112.4\degree$ in the quasi-ecliptic regime.  This $112\degree$ separation is not
a rounding artefact; it reflects the genuine angular divergence between the lunar
apogee and the descending node at this epoch.  No ambiguity in coordinate system,
ayanamsa choice, or computational precision could reduce this gap to zero.  A
practitioner who computes \zh{計都} as the apogee places it in the ninth house;
one who uses the node places it in the fifth.  These are qualitatively different
chart interpretations, and the choice between them is an astronomical question,
not a matter of lineage preference.

\subsection*{Implementation Note: The \texttt{stem-branch} Library}

The computations underlying Table~\ref{tab:four-regime} and
Figure~\ref{fig:ketu-period} are fully reproducible from the
\texttt{stem-branch} open-source library, which implements the
\zh{果老星宗} convention based on the scholarly consensus of Niu~(1994) and
Mak~(2014).  In the library's default configuration, \zh{計都} is assigned to
the osculating lunar apogee (apogee mode); the descending-node mode is
available as a configuration option to reproduce Qing-convention charts for
comparative purposes.

The library exposes both the \zh{似黃道} coordinate pipeline and the full
four-regime transformation chain, allowing researchers to generate positions
in any of the historically attested reference frames from a single API call.
All astronomical kernels delegate to the Swiss Ephemeris (JPL DE431 integration),
ensuring sub-arcsecond accuracy for dates within the historical scope of this
paper~\cite{mak2014pseudoplanets,niu1995ketu}.

\subsection*{Verification~3: Practical Significance of the \zh{似黃道}}

The quasi-ecliptic boundary computation whose derivation is presented in
Section~\ref{sec:ming} yields an additional quantitative result that can now be
stated precisely.  The maximum deviation of a \zh{似黃道} house boundary from
the corresponding equal-division boundary is $\pm 2.19\degree$.  The full
peak-to-trough range across the zodiac is therefore $4.38\degree$.

At Saturn's mean motion of $\approx 0.033\degree$/day, this $4.38\degree$ range
corresponds to approximately 133~days of positional ambiguity.  A Saturn that
lies within $2.19\degree$ of a house cusp in the \zh{似黃道} system could be
assigned to either of two adjacent houses depending on which version of the
coordinate boundaries the practitioner employs.  For Jupiter (mean motion
$\approx 0.083\degree$/day), the same angular tolerance corresponds to
$\approx 53$~days; for Mars ($\approx 0.52\degree$/day), to $\approx 8$~days.

This quantifies, for the first time, the practical significance of the
coordinate choice for historical chart interpretation.  The $4.38\degree$
figure is not a negligible rounding correction: at Saturn's pace it represents
four months of ambiguity in house placement, sufficient to alter the
qualitative reading of a nativity.  That Ming practitioners accepted this
margin of variation---rather than collapsing back to either pure ecliptic or
pure equatorial measurement---indicates that the \zh{似黃道} served a
principled purpose: it preserved continuity with inherited Sanskrit terminology
while accommodating the equatorial infrastructure of Chinese positional
astronomy, at the cost of a bounded but non-trivial coordinate
indeterminacy~\cite{mak2014pseudoplanets}.

% === §10 Discussion ===========================================================
\section{Discussion}\label{sec:discussion}

\subsection*{The Six Modes Validated Across Both Threads}

The evidence assembled in Sections~\ref{sec:tang}--\ref{sec:computation} allows
the six-mode framework introduced in \S\ref{sec:intro} to be evaluated against
two independent empirical threads: the coordinate reference frame and the
identity of \zh{計都}.  Both threads exhibit the same transformation patterns,
and the convergence of the two provides stronger support for the framework than
either thread could supply alone.

Three results are particularly important.  First, the modes are not sequential
developmental stages.  Mode~1 (transliteration) and Mode~2 (reidentification)
appear simultaneously in the Tang period
(\S\ref{sec:tang})~\cite{kotyk2018sinicization,mak2024persian}: the
\textit{Jiuzhi li} imported the Sanskrit terminology intact while the
\zh{七曜攘災訣} immediately began repurposing \textit{Ketu} as the apogee.
Transliteration and reidentification are logically independent operations, and
the historical record shows that they can---and did---proceed in parallel.

Second, Mode~3 (indigenous restructuring) and Mode~4 (hybridization) are
cumulative rather than mutually exclusive.  The Yuan equatorial shift
(\S\ref{sec:yuan}) created the structural precondition for the Ming synthesis:
without an established equatorial infrastructure, the \zh{似黃道} of
\S\ref{sec:ming} would have had no equatorial layer to hybridize with the
inherited ecliptic terminology~\cite{ho2003astrology}.  The modes can therefore
build on each other rather than simply succeeding one another.

Third, Mode~5 (foreign correction) can reset the system to an apparently
earlier state without reproducing that state exactly.  The Jesuit reform of
1645 (\S\ref{sec:jesuit}) returned the \zh{七政四餘} to an ecliptic-based
coordinate frame---ostensibly the same choice made in the Song
(\S\ref{sec:song})---but the new frame was tropical rather than sidereal.  A
reset is not a restoration.  This distinction has direct practical consequences:
the tropical and sidereal frames diverge by the ayanam\'{s}a ($\approx
24\degree$ at the present epoch), and the computational verification in
\S\ref{sec:computation} shows that the two Ketu conventions place the body in
houses that are separated by more than $112\degree$ for the test date.

Mode~6 (algorithmic fossilization, \S\ref{sec:fossil}) stands apart from the
other five in a theoretically important respect: it is the only passive mode.
Modes~1--5 all involve deliberate choices by identifiable historical actors---
translators, astrologers, reformers, missionaries.  Mode~6 requires no such
agency.  It operates through the inertia of implementation: once a convention
is encoded in software that is widely distributed, the convention persists until
an equally authoritative competing implementation displaces it.

\begin{table}[ht]
\centering
\caption{Six modes of cross-civilizational astronomical transformation.}
\label{tab:modes}
\small
\begin{tabular}{@{}llr@{}}
\toprule
Mode & Primary evidence & Section \\
\midrule
1. Transliteration & \zh{九執曆} (718~CE) & \S\ref{sec:tang} \\
2. Reidentification & Ketu: node $\to$ apogee & \S\ref{sec:tang} \\
3. Indigenous restructuring & Yuan equatorial shift & \S\ref{sec:yuan} \\
4. Hybridization & \zh{似黃道} system & \S\ref{sec:ming} \\
5. Foreign correction & Jesuit reform (1645) & \S\ref{sec:jesuit} \\
6. Algorithmic fossilization & Modern software & \S\ref{sec:fossil} \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{There and Back Again, But Not Quite}

The coordinate thread traces a trajectory that, viewed superficially, looks
circular.  The Indian original used an ecliptic sidereal frame
(\S\ref{sec:indian}).  The Tang transmission preserved that frame while
struggling to fit it into an equatorial Chinese context
(\S\ref{sec:tang}).  The Song codification (\S\ref{sec:song}) established
ecliptic sidereal measurement as the canonical baseline.  The Yuan reform
(\S\ref{sec:yuan}) abandoned ecliptic coordinates for equatorial ones.  The
Ming synthesis (\S\ref{sec:ming}) constructed the quasi-ecliptic hybrid, moving
back toward an ecliptic-like frame.  The Jesuit reform (\S\ref{sec:jesuit})
completed the return by adopting a fully ecliptic frame.

The system had apparently come home.

But the return destination was tropical rather than sidereal---the ``opposite
corner'' of the coordinate space from the Indian starting point.  Ecliptic
sidereal and ecliptic tropical agree on the plane of measurement but disagree
on the zero point.  The ayanam\'{s}a that separates them is not a small
correction but a displacement comparable to a full zodiacal sign.  A
practitioner who computes a chart in one system and reads it in the other will
find every planetary position shifted by approximately the same amount---a
systematic, uniform error that can be mistaken for a variant tradition.

This trajectory was not planned by any single actor.  The Tang translators
sought fidelity to their Sanskrit sources.  The Yuan astronomers sought
compatibility with the \textit{Shoushi li} equatorial infrastructure
\cite{sivin2009granting}.  The Ming compilers sought a vocabulary that their
practitioners could recognize.  The Jesuit reformers sought scientific
correctness by the standards of seventeenth-century European astronomy
\cite{liu2020missionary}.  Each decision was locally rational.  The cumulative
result---a system that is ecliptic but tropical where the original was ecliptic
but sidereal---emerged from centuries of independent choices without being
intended by any of them.

\subsection*{The Six Modes Beyond the China Case}

The Silk Road astronomical corridor carried knowledge in multiple directions
across multiple centuries, and the six modes described here are not specific to
the India--Persia--China pathway.  Two further transmission episodes illustrate
the framework's generality.

The translation movement of the House of Wisdom (Baghdad, 8th--9th centuries)
involved the transfer of Greek astronomical texts into Arabic.  The process
began with transliteration of technical terms (Mode~1)---Greek planetary names
were rendered phonetically before Arabic equivalents were established---and
proceeded through a phase of indigenous restructuring (Mode~3) in which Arab
astronomers embedded Greek parameters into the \textit{zij} tradition with its
own computational conventions~\cite{pingree1981jyotih}.  The Arabic
transformation of Ptolemaic astronomy was neither a passive copying nor a
wholesale replacement but a systematic reworking that parallels the Yuan
treatment of the Indian \zh{七政四餘}.

The \textit{Yavanaj\={a}taka} (2nd century CE) represents an earlier episode:
the transfer of Greek horoscopic astrology into Sanskrit.  Here, Mode~1 is
visible in the transliteration of Greek zodiacal terms (\textit{kri\'{y}\={a}}
for Aries, \textit{t\={a}vuri} for Taurus, and so on), while Mode~4 is visible
in the hybridization of the imported zodiacal framework with the existing
\textit{nakshatra} lunar mansion tradition, producing a system that is neither
purely Greek nor purely Vedic~\cite{pingree1981jyotih}.

These parallels suggest that the six-mode framework captures something general
about how astronomical knowledge moves across civilizational boundaries.  A
rigorous application to each episode---with the same dual-thread method applied
here---would require dedicated case studies; the present paper has established
the vocabulary and demonstrated the method.  What the parallels already confirm
is that the modes are not artefacts of the particular source and target cultures
under examination.

\subsection*{Limitations and Open Questions}

Several limitations constrain the current analysis.  Most importantly, the
characterizations of \zh{三辰通載} and \zh{鄭氏星案} in Sections~\ref{sec:song}
and \ref{sec:yuan} rely substantially on secondary analyses by
Niu~\cite{niu1995ketu,niu2010fourremains} and Mak~\cite{mak2014pseudoplanets}.
Critical editions of both texts would strengthen the argument in those
sections; the present paper has proceeded on the basis of the best available
scholarship, but primary philological work remains a prerequisite for
definitively establishing the coordinate conventions of the Song and Yuan
systems.

The \zh{似黃道} mathematical definition (\S\ref{sec:ming}) has been derived from
first principles and verified computationally (\S\ref{sec:computation}), but
cross-checking the derivation against attested historical computational
examples from the \zh{果老星宗} manuscript tradition would further consolidate
the result.  The computational agreement between the \texttt{stem-branch}
library and the independently reconstructed \zh{似黃道} boundary gives
confidence in the derivation, but it cannot substitute for documentary
confirmation.

The locus of the Ketu reidentification---whether the node-to-apogee shift
originated with Persian intermediaries or with early Chinese adapters---remains
genuinely open.  Mak's work on the Persian intermediary
\cite{mak2024persian} has narrowed the question without resolving it.  The
identification may have been polycentric, occurring independently in multiple
contexts, which would itself be a theoretically interesting finding.

Finally, this paper has examined one system (\zh{七政四餘}) within one
transmission corridor (India--Persia--China).  The extent to which the six-mode
framework generalizes to other astronomical traditions---Babylonian--Greek,
Arabic--Latin, Chinese--Korean--Japanese---remains to be tested empirically.

\subsection*{Relationship to the Companion Paper}

The companion paper ``When the Stars Disagree'' presents a computational error
taxonomy across five Chinese divination systems, documenting systematic
divergences between modern software implementations and their claimed historical
sources.  The coordinate-regime analysis developed here provides the historical
explanation for that paper's key finding in \S6.7: the tropical--sidereal
divergence that the computational survey identifies as the single largest source
of inter-implementation disagreement is not a technical error propagated by
inattentive developers.  It is the archaeological residue of Mode~5 (the Jesuit
foreign correction) overwriting an earlier Mode~3 convention (Yuan equatorial
sidereal) that had itself overwritten the original Mode~1 ecliptic sidereal
baseline.  The software disagreement is a stratified deposit; the present
paper supplies the stratigraphy.

% === §11 Conclusion ===========================================================
\section{Conclusion}\label{sec:conclusion}

This paper has traced two parallel threads---the coordinate reference frame and
the identity of the pseudo-planet \zh{計都}---through four dynasties and
thirteen centuries of the Chinese \zh{七政四餘} system.  The trajectory spans
four distinct coordinate regimes (ecliptic sidereal, equatorial sidereal,
quasi-ecliptic hybrid, ecliptic tropical) and two fully distinct astronomical
identities for \zh{計都} (descending lunar node; osculating lunar apogee).  Six
transformation modes account for every transition observed: transliteration,
reidentification, indigenous restructuring, hybridization, foreign correction,
and algorithmic fossilization.

The central finding is that cross-civilizational astronomical transmission is
not passive copying.  Each time the \zh{七政四餘} crossed a cultural or
institutional boundary---from Sanskrit into Chinese, from Tang court astronomy
into Song astrological practice, from the equatorial \textit{Shoushi} world
into Ming popular divination, from Chinese convention into Jesuit-reformed
state astronomy, from printed manuals into digital software---the transmitted
material was actively transformed.  The transformations were systematic: they
followed identifiable patterns that recur across the historical record and that
appear in other Silk Road transmission episodes.  But they were not
teleological: no single actor planned the trajectory, and the end state---a
tropical ecliptic frame with \zh{計都} as the lunar apogee---is not a
destination that any participant in the transmission history set out to reach.

This has an important consequence for how we interpret the fragmented,
contradictory landscape of modern \zh{七政四餘} software.  The confused
conventions, mismatched coordinate frames, and contested \zh{計都} toggles that
practitioners encounter today are not bugs awaiting correction.  They are
archaeological artifacts: each ``error'' preserves a trace of a legitimate
historical transformation.  Software that computes \zh{計都} as the descending
node is not wrong---it faithfully implements the Qing-era convention introduced
by Schall von Bell's reform.  Software that uses the lunar apogee is not
wrong---it faithfully implements the scholarly consensus established by
Niu~\cite{niu1995ketu} and Mak~\cite{mak2014pseudoplanets} regarding the
\zh{果老星宗} original.  The disagreement between them is real, but its source
is historical, not computational.

The six-mode framework developed here provides a vocabulary for analysing other
transmission episodes along the Silk Road astronomical corridor and beyond.  The
parallels noted in \S\ref{sec:discussion}---the House of Wisdom translations,
the \textit{Yavanaj\={a}taka} synthesis---suggest that the modes are general
features of how technical knowledge moves across civilizational boundaries,
rather than peculiarities of the China case.  Applying the framework rigorously
to further episodes, with the same dual-thread methodology used here, is a
natural direction for future work.

The deeper implication is historiographical.  Transformation is not a deviation
from transmission---it is the mechanism of transmission.  Every time a
practitioner in a new cultural context received the \zh{七政四餘}, they
encountered not a finished system but a set of tools and a set of problems, and
they rebuilt the system around the problems that mattered to them.  The system
that survived was not preserved; it was perpetually reconstructed.  Understanding
that history of reconstructions is not merely of antiquarian interest.  It is
the only reliable basis for deciding which convention any given historical source
actually implements---and therefore the only sound foundation for software that
faithfully computes what it claims to compute.

% === Bibliography =============================================================
\bibliographystyle{unsrtnat}
\bibliography{silk-road-astronomy}

\end{document}
